---
title: 'DSwR Workshop: Practice Session #1'
author: "Obryan Poyser"
date: "2/20/2019"
output:
    html_document:
        toc: true
        toc_float: true
        number_section: false
        theme: yeti
        highlight: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Hello!

This is a RMarkdown document, in here is possible to run the code and write text in a harmonic way. The learning curve is not that steep, but one has to get used to the commands and options. Nonetheless, I believe some things can be learned in parallel.

The chunk below is a mini R console, here is possible to execute any code as a R file. Take into account that I am talking about the .Rmd and not the html.

```{r}
log(100)
```

# Practice

## Q1 {.tabset .tabset-fade .tabset-pills}

### Statement
1. What about the mean of a vector composed by strings and numbers? Or a `NA` value?

### Solution

```{r}
vec_exc1 <- c(2:9, "a")
mean(vec_exc1)
vec_exc2 <- c(2:9, NA)
mean(vec_exc2)
mean(vec_exc2, na.rm = TRUE)
```

## Q2 {.tabset .tabset-fade .tabset-pills}

### Statement

1. Calculate the `mean()` statistic of a $3 \times 3$ matrix composed by the sequence 2 to 10 organized as a matrix. Is it the same as calculating a vector integrated by those numbers? 

### Solution

```{r}
mat_exc <- matrix(2:10, nrow = 3, ncol = 3)
(mat_exc_mean <- mean(mat_exc))

vec_exc <- 2:10
(vec_exc_mean <- mean(vec_exc))

mat_exc_mean==vec_exc_mean
```


